Structural replication of nanoporous media using procedural noise

### 2017-03-24 ms.tex
Structure of intro, still needs to be written
Rewrote introduction. Still missing references. Some changes to conclusions - revision of conclusion needed.

%During recent years, increase in computational power has allowed for a vast improvement  
\todo{Skriv en fin intro! Ogsaa om nanoporous media}
Properties of structural patterns that regularly occur in nature are often investigated either through direct observations or through simulations of theoretical models. Model simulations typically require a large computational framework to accurately reproduce the observed effects, while experimental data can be both expensive and hard to obtain, in addition to being riddled with artefacts and instrumental errors. While physical simulations yields statistically correct results, it is often hard to re-create features that are noticeable in experimental data, and therefore fit model parameters with observations. In this paper, we propose a new method for producing realistic-looking nanoporous media using only random noise fields, without the need for physics simulations. We do not claim that our method produces results that are more physically accurate than full model simulations, quite the contrary, we aim to show that it is possible to create fast, near-accurate and arbitrary large data sets of nanoporous media that are statistically similar to physics-based simulations with only the use of random noise fields.

Procedural generation of physical structures have several advantages over conventional simulations. For example, when producing simulations of nanoporous silica, we are required to calculate and propagate the physical properties of each particle for every time-step throughout the simulation, using a computational framework such as LAMMPS \cite{plimpton1995fast}. This is both time-consuming, and the computational cost can sometimes be be quite large. In addition, the simulation is constrained to a box, so particles are not allowed to move outside the defined boundaries. On the other hand, using procedural noise to carve out nanoporous media is both extremely fast and computationally cost-efficient, in addition to being continuous on all scales.

\todo{Nevn finite size effects}
%In addition, by creating a model that yields properties that are statistically consistent with observations, we are ar

Since we aim to investigate whether procedural methods are suitable for producing statistically correct 3D physical structures, we choose to focus our attention on silica-based nanoporous media. In a sense, we are combining well-known methods from computer games and real-time graphics-oriented software with experimental observations and simulations. Procedural methods for automated generation of computer game content have been around since the early 1980's, and in those early days algorithms were less advanced, often yielding plain, simple geometrical structures. Nowadays, almost every single computer game produced will contain some level of automated generated content, from textures to level design to landscapes and gameplay.

We do not claim that procedural noise methods completely reproduces all structural properties of nanoporous silica, but show that our approach can be used to efficiently create materials that are structurally similar to simulations. Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the radial distribution function $g(r)$ (RDF). 
\todo{Nevne at kjemi naturligvis ikke vil være riktig siden antall naboer per atom ikke er riktig. }
%\section{Method}
\section{Procedural content generation}
By procedural content generation, we generally mean a mathematical algorithm that uses limited input data to create complex, sometimes infinite repeating patterns. A firm example would be fractals, such as the Julia/Mandelbrot [ref Mandelbrot] set, where one tests whether the complex quadratic recurrence equation $z_{n+1} = z_n^2 +C$ for $z \in \mathbb C$ converges for each position in a 2D complex plane. With this seemingly simple equation, the level of complexity generated on the boundary between convergence and divergence is infinite, with self-similar patterns repeating themselves to arbitrary detail.
\begin{figure}
\includegraphics[width=.5\textwidth]{orphancity.png}
\caption{Example of city building using procedural methods: Landscape, city infrastructure, buildings and all textures are all automatically generated using procedural methods. Created by one of the authors. \todo{Denne er jo med bare for moro skyld, boer jo fjernes} }
\label{fig:orphancity}
\end{figure}

@@

\subsection{Some notes on the results}
It is worth to note that since we are chiefly interested in obtaining a set of parameters that reproduces a statistical correct representation of the input data, we are relatively free to choose parameters as desired within the posterior distribution width of the parameter. In the previous analysis, we selected a best fit threshold value of $\tau=-0.2$, but from the width of the likelihood it is clear that nearby parameters would also be suitable. However, the surface area is strongly correlated with this value, so even slight deviations (e.g. $\Delta \tau = 0.05$) yields a relatively large change in surface area. In addition, we do not claim that this model completely reproduces the pore structure observed in nanoporous simulations, as is evident from the RDF shown in figure \ref{fig:gofr1}, but hope to discover a better model that will produce a $g(r)$ that is more in accordance with the data.

@@

\section{Conclusion}
\todo{konklusjon maa renskrives}
In this paper, we have proposed a novel method for generating a simple kind of nanoporous media using procedural methods, namely simplex noise. Procedural methods, albeit having no root in physics, have some advatages over prevalent methods: it computationally efficient, has low memory footprints and is only limited in terms of detail by the model in question. With detail, we mean that is possible to create simulations of arbitrary scales and resolution. We have also developed a framework for both producing such materials and also estimating model parameters given data. We have shown that it is possible through a full likelihood analysis to successfully reproduce the input parameters of procedurally simulated data. In addition, we performed an analysis on a simulated $\SI{159}{\angstrom^{3}}$ data set of SiO$_2$, and explain that while porosity and surface area corresponds well between the model and data, the $g(r)$ shows that the current model does not completely reproduce correct data. However, we do not claim that the current model should do so yet, since we have chosen the simplest kind of summation with regards to procedural noise modes.

>>>>

The rapid development of new experimental and observational methods to determine the three-dimensional structure of many types of porous media calls for new statistical tools and methods to characterize complex porous media and to generate synthetic media based on the observed structures. New experimental methods allow for studies of three-dimensional structures from nano-meter to meter scales. Destructive techniques such as focused ion beam scanning electron microscopy (FIB-SEM) are routinely used to characterize geologically relevant structures in shales and carbonate rocks with resolutions down to tens of nanometers \cite{XX}. High-energy X-ray tomography methods using high-energy beams at international research facilites or in-house scanners open for three-dimensional mapping of mineral structures from sub-micrometer to millimeter scales \cite{XX}. Thus new experimental techniques provide three-dimensional voxelized images that can form the basis of analysis or modeling. However, in many cases we can only scan a few examples experimentally, but need to perform simulations over many statistically similar realizations to find effective properties or to populate upscaling models. Thus the need for methods to robustly characterize three-dimensional porous structures and to generate synthetic structures with similar statistical properties.

The field of synthetic rocks is a large field both scientifically and technologically \cite{}. Methods to generate synthentic rock samples from experimental characterizations may be based on purely statistical or geometrical models, models of physical processes that generate the samples, or combinations of both. For small-scale porous media, physical models are often used to generate initial grain packings \cite{}, then further mechano-chemical and fluid-mediated processes such as precipitation-dissoluton processes and pressure solution \cite{}, or fracture and plastic deformation processes are added to include effects of long-term compaction \cite{}. On large scales, reservoir models are populated by statistical models for lithological and geomorphological processes \cite{}, such as river fan formation to generate particular grain size distributions \cite{}. More statistical approaches include variations of percolation models \cite{sahimi} or network models \cite{} to characterize and generate uncorrelated or correlated porous media. In addition, the high resolution of observational methods allow three-dimensional scans to form the basis for atomic-scale modeling using molecular modeling methods such as molecular dynamics or Monte Carlo methods to study the dynamics of nanoporous media. But then procedural or physical models are needed to fill in the atomic scale details, to build many statistical realizations of nanoporous media with atomic-scale details, and to extrapolate to larger scale models.

These physically based models may not always include all relevant effects, can be quite computationally expensive when used to generate many samples, and cannot easily be extended beyond the initial size of the simulated systems. For example, to prepare a simulation of nanoporous silica, we need to carry out a sequence of large scale simulations in molecular dynamics, which is time consuming, computationally expensive and the results canno be exteded beyond the intial size of the simulated systems without performing another simulation \cite{Adarsh}. On the other hand, the statistical models often do not include all the correlations induced by physical processes or have limited number of tunable parameters. In particular, statistical models have challenges in reproducing atomic-scale structures of nanoporous media.

Meanwhile, there has been a recent growth in advanced random noise models that open for new ways to generate structured noise with statistical similarities to real samples, such as the use of Perlin noise. Such noise models have been used to model systems with complex statistical correlations such as in (examples from Nicloaas galaxy/astrophysics applications), and show great promise in being adaptable to many different physical systems. Also, such noise models are continuous and extendable, so that arbitrary large systems or systems with arbitrary high precision can be modeled. While these models show great promise in many fields \cite{}, they have not yet been systematically applied to characterize or generate nanoporous media with statistical properties corresponding to that of real systems.

In this paper, we propose a new method for producing realistic-looking nanoporous media using only random noise fields, without the need for physics simulations. We provide a systematic study of the use of procedural noise-based models to characterize and generate three-dimensional nanoporous media. Our studies will be based on silica-based nanoporous media, for which we have good atomic scale models and physics-based methods to generate porous media as a sequence of simulated processes, and we therefore can generate large-scale samples where we have full knowledge of all aspects of the material. These simulated systems are then characterized using the radial distribution function, $g(r)$, statistical methods are generated with similar properties, and compared with the simulated systems. Using these approaches we find that procedural noise methods can be tuned to efficiently create materials that are structurally similar to simulationed materials.

\section{Procedural content generation}
By procedural content generation, we generally mean a mathematical algorithm that uses limited input data to create complex, sometimes infinite repeating patterns. A firm example would be fractals, such as the Julia/Mandelbrot [ref Mandelbrot] set, where one tests whether the complex quadratic recurrence equation $z_{n+1} = z_n^2 +C$ for $z \in \mathbb C$ converges for each position in a 2D complex plane. With this seemingly simple equation, the level of complexity generated on the boundary between convergence and divergence is infinite, with self-similar patterns repeating themselves to arbitrary detail.
\begin{figure}
\includegraphics[width=.5\textwidth]{orphancity.png}
\caption{Example of city building using procedural methods: Landscape, city infrastructure, buildings and all textures are all automatically generated using procedural methods. Created by one of the authors. \todo{Denne er jo med bare for moro skyld, boer jo fjernes} }
\label{fig:orphancity}
\end{figure}

@@

\subsection{Discussion}
It is worth to note that since we are chiefly interested in obtaining a set of parameters that reproduces a statistical correct representation of the input data, we are relatively free to choose parameters as desired within the posterior distribution width of the parameter. In the previous analysis, we selected a best fit threshold value of $\tau=-0.2$, but from the width of the likelihood it is clear that nearby parameters would also be suitable. However, the surface area is strongly correlated with this value, so even slight deviations (e.g. $\Delta \tau = 0.05$) yields a relatively large change in surface area. In addition, we do not claim that this model completely reproduces the pore structure observed in nanoporous simulations, as is evident from the RDF shown in figure \ref{fig:gofr1}, but hope to discover a better model that will produce a $g(r)$ that is more in accordance with the data.

@@

\section{Conclusion}
\todo{konklusjon maa renskrives}

In this paper, we have proposed a novel method for generating a nanoporous media using procedural methods in the form of simplex noise. Procedural methods, albeit having no root in physics, have some advatages over prevalent methods: they are computationally efficient, have low memory footprints and are only limited in terms of detail by the model in question. In particular, it is possible to create simulations of arbitrary scales and resolution, indeed with any level of detail.

\todo{AMS: This is a bit weak currently. State first what you have succeeded in doing. Then state the possible weaknesses of the model.}
We have developed a framework to generate nanoporous media using molecular dynamics simulations and to estimate model parameters given such data. We have shown that it is possible through a full likelihood analysis to successfully reproduce the input parameters of procedurally simulated data. In addition, we performed an analysis on a simulated $\SI{159}{\angstrom^{3}}$ data set of SiO$_2$, and explain that while porosity and surface area correspond well between the model and data, analysis of $g(r)$ shows that the current model reproduces the main features of the material, but that there are some deviations. We find the correspondence strong, which demonstrates the proof-of-principle of the use of procedural noise methods, but we realize that even better correspondence may be achieved by including other summations in the procedural noise modes.
